---
title: "Module2"
output:
  word_document: default
  html_document: default
---

# Introduction  

In this file we are downloading a file from the web, importing it, printing out descriptive staistics, transforming a variable, plotting a quantitative variable, creating a scatterpliot, and uploding to Git 

```{r warning = FALSE}
library(tidyverse)
library(datasets)
```
## Get data and print out summary statistics

We called the datasets library so we can use the famous Iris dataset directly. Lets assign the table to one we name, and print out some summary statistics

```{r}
iris_table <-iris

summary(iris_table)
```

#Transform a variable
Here we will use dplyr to call the dataframe, then multiply a column by 1.1, and set it back to itself. The summary stats should change for that column

```{r}
iris_table <- iris_table %>%
  mutate(Petal.Length = Petal.Length*1.1)

summary(iris_table)
```

#Plot a quantitative variable
We will now plot a quantitative variable. We will make a box plot of Sepal Length by Species

```{r}
plot(iris_table$Species, iris_table$Sepal.Length, main="Sepal Length by Species", xlab="Species", ylab="Sepal Length")
```

#Create a scatterplot
We will now create a scatterplot. We will plot Sepal Length by Petal Length

```{r}
plot(iris_table$Sepal.Length, iris_table$Petal.Length, main="Sepal Length vs Petal Length", xlab="Sepal Length", ylab="Petal Length")
```



